Hypertension and microbiome

### 2019-01-25 articletwo.Rmd
Minor

```{r variables, warning = FALSE}
data("phfinrisk_metadatadesc", package="finriskmetagcommon")
names.dset <- filter(phfinrisk_metadatadesc,
                     Ignored.Covariate.in.Cross.Sectional.Analysis.Aaro20181115==0)
names.dset <- bind_rows(names.dset,
                        data.frame(Covariate = c("MAP", "PULSEPRESSURE", "HYPERTENSION", "SEX"),
                                   Category = rep("Physical", 4),
                                   Name =c( "Mean artrerial pressure", "Pulse pressure", "Hypertension", "Female"),
                                   Desc = c("Mean arterail presure, 2./3.*DIASM + 1./3.*SYSTM)",
                                            "Pulse pressure, SYSTM - DIASM",
                                            "Hypertension, SYSTM >=140 or DIAS >= 90 or BP mediaction",
                                            "Sex is female, True for female, False for male")))
```

@@

@@

```{r alphabeta definitions}
dset.alpha <- meta.merge.alphadiversity(pseq.species)
alphadiversity <- list(
    min = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
    max = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon"))
betadiversity  <- readRDS("rds/adonis.species.rds")
diversities <- mergediversities(alphadiversity, betadiversity)

@@

```{r pca}
pca.prcomp <- readRDS("rds/pca.prcomp.rds")
pca.eigs <- pca.prcomp$sdev^2
pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)

@@

>>>>

```{r variables, warning = FALSE}
names.dset <- getdescriptions()
```

@@

## Models

```{r matrix calculation}
if (!file.exists("rds/bray.dist.m.species.rds") || !file.exists("rds/bray.dist.m.genus.rds")) {
    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
} else {
    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
}
```

 ```{r adonis calculation}
 if (!file.exists("rds/adonis.species.rds")) {
     adonis.species = list()
     adonis.species$min <- calculateadonis(dset = meta(pseq.species),
                                           matrix = bray.dist.m.species,
                                           covariates = c("BL_AGE", "SEX"),
                                           npermutations = 99)
     adonis.species$max <- calculateadonis(dset = meta(pseq.species),
                                           matrix = bray.dist.m.species,
                                           covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
                                                          "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                                           npermutations = 99)
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
    } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
}
```

```{r pca}
if (!file.exists("rds/pca.prcomp.rds")) {
    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
    pca.prcomp <- stats::prcomp(pca.abundances)
    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
} else {
    pca.prcomp <- readRDS("rds/pca.prcomp.rds")
}
```

```{r run maaslin, results = 'hide'} 
maaslin.force.vars <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")

prepare.maaslin(pseq.genus,
                 tsvfile = "maaslin_core_merged_phfinrisk_genus.tsv",
                conffile = "maaslin_core_config.config",
                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars),
                core = TRUE,
                readpclrows = "Halapricum.Archaea-")

Maaslin("maaslin_core_merged_phfinrisk_genus.tsv",
        "maaslin_core",
        strInputConfig = "maaslin_core_config.config",
        strModelSelection="none",
        dSignificanceLevel = 1.0, 
        fAllvAll = TRUE,
        strForcedPredictors = maaslin.force.vars)
```

```{r lactobacillus scatterplot,  results = 'hide'}
pseq.lacto  <- subset_taxa(pseq.species, Genus=="Lactobacillus")
maaslin.force.vars.lacto <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")

prepare.maaslin(pseq.lacto,
                 tsvfile = "maaslin_lacto_merged_phfinrisk_genus.tsv",
                conffile = "maaslin_lacto_config.config",
                includedvars = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION", maaslin.force.vars.lacto),
                readpclrows = "Lactobacillus_delbrueckii.Bacteria,Lactobacillus_ruminis.Bacteria,Lactobacillus_acidophilus.Bacteria,Lactobacillus_fermentum.Bacteria,Lactobacillus_paracasei.Bacteria,Lactobacillus_salivarius.Bacteria")

Maaslin("maaslin_lacto_merged_phfinrisk_genus.tsv",
        "maaslin_lacto",
        strInputConfig = "maaslin_lacto_config.config",
        strModelSelection="none",
        dSignificanceLevel = 1.0, 
        fAllvAll = TRUE,
        strForcedPredictors = maaslin.force.vars.lacto)
```

@@

```{r alphabeta definitions}
dset.alpha <- meta.merge.alphadiversity(pseq.species)
alphadiversity <- list(
    min = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
    max = calculateglm(dset.alpha, modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09  + diversities_shannon"))
diversities <- mergediversities(alphadiversity, adonis.species)

@@

```{r pca}
pca.eigs <- pca.prcomp$sdev^2
pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)

@@

### Lactobacillus

```{r Reading maaslin lacto output}
maas.lacto.results.df.signf <- read.table("maaslin_lacto/maaslin_lacto_merged_phfinrisk_genus.txt", header=TRUE, sep='\t') %>%
    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
                  beta = Coefficient) %>%
   dplyr::filter(Qval < 0.05)

maas.lacto.signf.features <- maas.results.df.signf %>%
    pull(Feature) %>%
    as.character() %>%
    gsub("_", ".", .)

df.lacto.merged <-  merge.pheno.abu(pseq.lacto)
```
